Python as a machine learning language: experiences from the scikit-learn project

### 2013-06-15 paper.tex
remove remark
implementation

\textcolor{red}{TODO: Mention that pipelines can also be grid-searched.}

@@

\noindent \textcolor{red}{TODO: Mention small set of dependencies.}\\
\textcolor{red}{TODO: elaborate on use of mixins and duck typing}

scikit-learn is primarily implemented in Python and Cython
\citep{behnel2011cython},
a language that extends Python with static typing
and a compiler that produces C extension modules
for the Python runtime system.
In addition, it includes (modified versions of)
the \textsf{LIBSVM} and \textsf{LIBLINEAR} libraries
used for training support vector machines
and logistic regression models \citep{chang2011libsvm, fan2008}.
These libraries are both written in C{}\verb!++!
and wrapped using Cython modules.

As much as possible is implemented in ``pure'' Python
with NumPy for numerical work,
to avoid the compile-edit-debug cycle needed for Cython
and to keep the codebase readable for as large an audience as possible.
However, not all algorithms are easily and efficiently expressed
as NumPy vector operations
(among other factors, because they often have linear space requirements).
Algorithms that fall into this category have to be implemented in Cython
to obtain decent performance and scalability.
Examples include $k$-means, stochastic gradient descent for linear models
and some graph-based clustering algorithms.

While object-oriented techniques and design patterns
are used throughout scikit-learn,
they mostly serve to facilitate code reuse and are not considered a project goal.
Where possible, the Python principle of ``duck typing'' is exploited
to simplify the implementation and skip the introduction of superfluous classes.
This allows for extensibility and flexibility at the same time:
user-defined estimators that follow scikit-learn conventions
should be usable in pipelines and other composite objects
without them actually inheriting from scikit-learn base classes.

%\section{Weaknesses of the NumPy/SciPy environment}

\textcolor{red}{TODO: The following corresponds to the old weaknesses section. It should be properly rewritten.}

>>>>

@@

Our implementation guidelines emphasize on writing short and readable code,
easily maintainable by external contributors, while being as efficient as
possible.  Whenever practicable, algorithms implemented in \sklearn are thus
directly written in pure Python. For intensive numerical computations, the
library mostly relies on NumPy vector operations, which allows for the code to
remain concise and readable while being efficient. For all algorithms that
cannot be easily and efficiently expressed as NumPy operations, we rely on
Cython \citep{behnel2011cython}  for writing the critical parts and achieving
decent performance and scalability. Cython is a compiled programming language
that extends Python with static typing. It produces efficient C extension
modules that are directly importable from the Python runtime system. Examples of
algorithms written in Cython include stochastic gradient descent for linear
models, some graph-based clustering algorithms or decision trees. To facilitate
the adoption of the library, we also limit to a bare minimum the set of external
depencies. As such, NumPy and SciPy are actually the only two required
dependencies for the library to fully work. Some modules may rely on other
external libraries, but then we tend to prefer to directly ship them within the
codebase rather than adding new dependencies. In particular, the codebase ships
modified versions of the \textsf{LIBSVM} and \textsf{LIBLINEAR} libraries
\citep{chang2011libsvm,fan2008}, both written in C{}\verb!++! and wrapped using
Cython modules.

% scikit-learn is primarily implemented in Python and Cython
% ,
% a language that extends Python with static typing
% and a compiler that produces C extension modules
% for the Python runtime system.
% In addition, it includes (modified versions of)
% the \textsf{LIBSVM} and \textsf{LIBLINEAR} libraries
% used for training support vector machines
% and logistic regression models \citep{chang2011libsvm, fan2008}.
% These libraries are both written in C{}\verb!++!
% and wrapped using Cython modules.

% As much as possible is implemented in ``pure'' Python
% with NumPy for numerical work,
% to avoid the compile-edit-debug cycle needed for Cython
% and to keep the codebase readable for as large an audience as possible.
% However, not all algorithms are easily and efficiently expressed
% as NumPy vector operations
% (among other factors, because they often have linear space requirements).
% Algorithms that fall into this category have to be implemented in Cython
% to obtain decent performance and scalability.
% Examples include $k$-means, stochastic gradient descent for linear models
% and some graph-based clustering algorithms.

% While object-oriented techniques and design patterns
% are used throughout scikit-learn,
% they mostly serve to facilitate code reuse and are not considered a project goal.
% Where possible, the Python principle of ``duck typing'' is exploited
% to simplify the implementation and skip the introduction of superfluous classes.
% This allows for extensibility and flexibility at the same time:
% user-defined estimators that follow scikit-learn conventions
% should be usable in pipelines and other composite objects
% without them actually inheriting from scikit-learn base classes.
